Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-12-18 paper/main.tex
reflect bug fixes from the code in the rules
continue updating the paper

\iffull{
\begin{figure*}
\begin{mathpar}
%
\Rule[leftskip=.4cm,rightskip=.3cm]{Q:Skip}
{ }
{ B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) }
%
\and
\Rule{Q:Break}
{ }
{ (\Gamma, Q_B); R; (\Gamma, Q_B) \vdash \code{break} \dashv (\Gamma', Q') }
%
\and
\Rule[leftskip=.3cm,rightskip=.4cm]{Q:Tick}
{ n < 0 \implies Q \ge 0 }
{ B; R; (\Gamma, Q) \vdash \code{tick}(n) \dashv (\Gamma, Q{-}n) }
%
\and
\Rule{Q:Return}
{ P = Q_R[\Vret/x]
\\ \Gamma = \Gamma_R[\Vret/x]
\\ \forall i \in \dom{P} . \, p_i = q_i
}
{ B; (\Gamma_R,Q_R); (\Gamma, Q) \vdash \code{return}~x \dashv (\Gamma',Q') }
%
\and \Rule{Q:Update}
{ q'_{xy}, q'_{yx} \in \Q
\\ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ B; R; (\Gamma[x/y], Q {+} M_u {+} M_e(y)) \vdash x \gets y \dashv (\Gamma, Q') }
%
\and \Rule[leftskip=.1cm,rightskip=.1cm]{Q:Loop}
{ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q) }
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\and \Rule{Q:Inc}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{ux}, -q_{xu})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{xu}, -q_{ux})
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:Dec}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{xu}, -q_{ux})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{ux}, -q_{xu})
}
{ \textstyle
  B; R; (\Gamma[x/x {-} y], Q) \vdash x \gets x - y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:If}
{ B; R; (\Gamma \land e, Q) \vdash S_1 \dashv (\Gamma',Q')
\\\\ B; R; (\Gamma \land \neg e, Q) \vdash S_2 \dashv (\Gamma',Q')
}
{ B; R; (\Gamma, Q) \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv (\Gamma',Q') }
%
\and\Rule{Q:Seq}
{  B; R; (\Gamma, Q) \vdash S_1 \dashv (\Gamma', Q')
\\\\ B; R; (\Gamma', Q') \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and\Rule{Q:Call}
{ (\Gamma_f, Q_f,  \Gamma_f', Q_f') \in \Delta(f) \!
\\ \text{Loc} = \text{Locals}(Q) \!
\\ \forall i \neq j . \, x_i \neq x_j \!
\\ c \in \Qplusz \!
\\ Q = P + S \!
\\ Q' = P' + S
\\ U = Q_f[\Vargs / \vec x]
\\ U' = Q'_f[\Vret/r]
\\ \forall i \in \dom{U} . \, p_i = u_i
\\ \forall i \in \dom{U'} .\, p'_i = u'_i
\\ \forall i \not\in \dom{U'} .\, p'_i = 0
\\ \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ B; R; (\Gamma_f[\Vargs / \vec x] \land \Gamma_{\text{Loc}}, Q {+} c)
  \vdash r \gets f(\vec x) \dashv
  (\Gamma_f'[\Vret/r] \land \Gamma_{\text{Loc}}, Q' {+} c)
}
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Assert}
{ }
{ B; R; (\Gamma, Q) \vdash \code{assert}~e \dashv (\Gamma \land e, Q) }
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Extend}
{ \fenv f = (\vec y, S_f)
\\ Q' \ge 0
\\ Q'_f \ge 0
\\\\ B; (\Gamma_f', Q'_f); (\Gamma_f[\Vargs/\vec y], Q_f[\Vargs/ \vec y] ) \vdash S_f \dashv (\Gamma', Q')
}
{ (\Gamma_f, Q_f, \Gamma_f', Q_f') \in \Delta(f) }
%
\and \Rule{Q:Weak}
{  \!\! \Gamma_1 \models \Gamma_2
\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ \Gamma'_2 \models \Gamma'_1
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy} {\in} \N \, . \, \Gamma \models l_{xy} \le |\inter x y| \}
\\ \mathcal U = \{ xy \mid \exists u_{xy}{\in} \N \, . \, \Gamma \models |\inter x y| \le u_{xy} \}
\\ \forall i.\, p_i, r_i \in \Qplusz
\\\\ \forall i \in \mathcal U.\, q'_i \ge q_i - r_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U {\cup} \mathcal L {\cup} \{0\} .\, q'_i \ge q_i
\\ q'_0 \geq q_0 {\textstyle + \sum_{i \in \mathcal U} u_i r_i - \sum_{i \in \mathcal L} l_i p_i}
}
{ Q' \succeq_\Gamma Q }
\end{mathpar}
\caption{Inference rules of the quantitative analysis.}
\label{fig:auto}
\end{figure*}
}{}

\ifshort{
                    \jan{XXX PLEASE INSERT FIGURE HERE XXX}
}{}

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x - 10 \geq 0)~\{x=x{-}10; tick(5);\}~else~break;}}}
\begin{figure*}[t]
\vspace{-.5cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (x {<} 10, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (\cdot, P^\text{de})
          }
        }
        { (x {<} 10, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (x {<} 10, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (x {<} 10, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (x {<} 10, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (x {<} 10, B^\text{br}); (x {<} 10,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (x {<} 10, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (x {<} 10, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (x {<} 10, P)
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(\cdot)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced $\toolname$. The constraints are resolved by an off-the-shelf LP solver.}
\label{fig:derivation}
\end{figure*}

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context that is
used to gather information on interval sizes.  For example, $P^\text{de} \succeq_{(\cdot)} P^\text{we} \equiv$
\vspace{-.3cm}
{\small
  \begin{align*}
 p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0.10}
\land  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10} \\
\land  \forall (\alpha, \beta) \neq (0,10)  . \, p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;\;\; .
  \end{align*}
}
\vspace{-.8cm}

@@

\vspace{-.24cm}

@@

\iffull{}{ The main parts of the soundness proof are formalized with
  Coq and available for download. The full definitions of the cost
  semantics and the quantitative Hoare logic, and more details on the
  soundness proof can be found in the extended version of this
  article.}

@@

The small-step semantics is standard, except that it tracks
the resource consumption of a program.  The semantics is parametric in
the resource of interest for the user of our system.  We achieve this
independence by parameterizing evaluations with a resource \emph{metric}
$M$; a tuple of rational numbers and two maps.  Each of these
parameters indicates the amount of resource consumed by a corresponding
step in the semantics.  Resources can be released by using a negative
cost. %  The metric used in the implementation can also
% depend on other information that is statically available such as the
% name of the called function or the number of arguments.
Two sample rules for update and tick follow.
\begin{mathpar}
\Rule[rightskip=0.3cm]{U}
{ \state' = \state[x \mapsto \sem{e}_\state] }
{ (\state, x \gets e, \cont, \cost)
\rightarrow\\\\
(\state', \code{skip}, \cont, \cost {-} M_u {-} M_e(e))
}
\and
\Rule[leftskip=0.3cm,rightskip=0.2cm]{T}
{ }
{ (\state, \code{tick}(n), \cont, \cost)
\rightarrow\\\\
(\state, \code{skip}, \cont, \cost {-} M_t(n))
}
\end{mathpar}
The rules have as implicit side condition that $\cost$ is
non-negative.  This makes it possible to detect a resource crash as a
stuck configuration where $\cost<0$.

@@

We define quantitative Hoare triples
as $B;R\vdash\htriple Q S {Q'}$ where $B$, $R$, $Q$, and $Q'$ are
maps from program states to an element of $\Qplusz\cup\{\infty\}$ that represents
an amount of resources available.  The assertions $B$ and $R$ are
postconditions for the case in which the block $S$ exits by a \code{break}
or \code{return} statement.  Additionally, $R$ depends on the return
value of the current function.  The meaning of the triple $\htriple Q
S {Q'}$ is as follows: If $S$ is executed with starting state $\state$,
the empty continuation, and at least $Q(\state)$ resource units available
then the evaluation does not run out of resources and there are at
least $Q(\state')$ resources left if the evaluation terminates in
$\state'$.  The logic rules are similar to the ones in previous
work and generalized to account for the cost introduced by  our
cost-aware semantics.

@@

The translation of an assertion $(\Gamma, Q)$ in the automatic
analysis is defined by
$$
  \tr(\Gamma, Q) := \lambda\state .\, (\Gamma(\state)) + \Phi_Q(\state),
$$
where we write $\Phi_Q$ for the unique linear potential function
defined by the quantitative annotation $Q$.  The logical context $\Gamma$
is implicitly lifted to a quantitative assertion by mapping a state
$\state$ to 0 if $\Gamma(\state)$ holds and to $\infty$ otherwise.
\iffull{We also need to
translate the assumptions in function contexts of the automatic
analysis.  We define $\tr(\Gamma_f, Q_f, \Gamma'_f, Q'_f) :=
\forall z \, \vec v \, v . $
%
$$
(
  \lambda \vec v \, \state .\, \tr(\Gamma_f, Q_f)(\state[\Vargs {\mapsto} \vec v]),
  \lambda v \, \state .\, \tr(\Gamma'_f, Q'_f)(\state[\Vret {\mapsto} v])
  )
$$}{}
%
\iffull{These definitions let}{This lets} us translate the judgement
$J := B; R; P \vdash S \dashv P'$\iffull{ in the context $\Delta$ by $\tr(J) :=$}{ by}
\iffull{$$
  (\lambda f .\, \tr(\Delta(f))); \tr(B); \tr(R) \vdash
    \htriple{\tr(P)}{S}{\tr(P')}.
$$}{%
$$
  \tr(J) := \tr(B); \tr(R) \vdash \htriple{\tr(P)} S {\tr(P')}.
$$}%
\noindent{}The soundness of the automatic analysis can now be stated
formally with the following theorem.
%
\begin{theorem}[Soundness of the automatic analysis]
  If $J$ is a judgement derived by the automatic analysis,
  then $\tr(J)$ is a quantitative Hoare triple derivable
  in the quantitative logic.
\end{theorem}
%
\noindent
The proof of this theorem is constructive and maps each rule
of the automatic analysis directly to its counterpart in the
quantitative logic.  The most tricky parts are the translations of the
rules for increments and decrements and the rule {\sc Q:Weak} for
weakening because they make essential use of the algebraic properties
of our linear potential functions.
\iffull{ For instance we prove the following lemma to show the soundness
of the translation of {\sc Q:Weak}.

>>>>

\iffull{
\begin{figure*}
\begin{mathpar}
%
\Rule[leftskip=.4cm,rightskip=.3cm]{Q:Skip}
{ }
{ B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) }
%
\and
\Rule{Q:Break}
{ }
{ (\Gamma, Q_B); R; (\Gamma, Q_B) \vdash \code{break} \dashv (\Gamma', Q') }
%
\and
\Rule[leftskip=.3cm,rightskip=.4cm]{Q:Tick}
{ n < 0 \implies Q \ge 0 }
{ B; R; (\Gamma, Q) \vdash \code{tick}(n) \dashv (\Gamma, Q{-}n) }
%
\and
\Rule{Q:Return}
{ P = Q_R[\Vret/x]
\\ \Gamma = \Gamma_R[\Vret/x]
\\ \forall i \in \dom{P} . \, p_i = q_i
}
{ B; (\Gamma_R,Q_R); (\Gamma, Q) \vdash \code{return}~x \dashv (\Gamma',Q') }
%
\and \Rule{Q:Update}
{ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ B; R; (\Gamma[x/y], Q {+} M_u {+} M_e(y)) \vdash x \gets y \dashv (\Gamma, Q') }
%
\and \Rule[leftskip=.1cm,rightskip=.1cm]{Q:Loop}
{ Q \ge Q'
\\ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q)
}
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\and \Rule{Q:Inc}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{ux}, -q_{xu})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{xu}, -q_{ux})
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:Dec}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{xu}, -q_{ux})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{ux}, -q_{xu})
}
{ \textstyle
  B; R; (\Gamma[x/x {-} y], Q) \vdash x \gets x - y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:If}
{ B; R; (\Gamma \land e, Q) \vdash S_1 \dashv (\Gamma',Q')
\\\\ B; R; (\Gamma \land \neg e, Q) \vdash S_2 \dashv (\Gamma',Q')
}
{ B; R; (\Gamma, Q) \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv (\Gamma',Q') }
%
\and\Rule{Q:Seq}
{  B; R; (\Gamma, Q) \vdash S_1 \dashv (\Gamma', Q')
\\\\ B; R; (\Gamma', Q') \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and\Rule{Q:Call}
{ (\Gamma_f, Q_f,  \Gamma_f', Q_f') \in \Delta(f) \!
\\ \text{Loc} = \text{Locals}(Q) \!
\\ \forall i \neq j . \, x_i \neq x_j \!
\\ c \in \Qplusz \!
\\ Q = P + S \!
\\ Q' = P' + S
\\ U = Q_f[\Vargs / \vec x]
\\ U' = Q'_f[\Vret/r]
\\ \forall i \in \dom{U} . \, p_i = u_i
\\ \forall i \in \dom{U'} .\, p'_i = u'_i
\\ \forall i \not\in \dom{U'} .\, p'_i = 0
\\ \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ B; R; (\Gamma_f[\Vargs / \vec x] \land \Gamma_{\text{Loc}}, Q {+} c)
  \vdash r \gets f(\vec x) \dashv
  (\Gamma_f'[\Vret/r] \land \Gamma_{\text{Loc}}, Q' {+} c)
}
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Assert}
{ }
{ B; R; (\Gamma, Q) \vdash \code{assert}~e \dashv (\Gamma \land e, Q) }
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Extend}
{ \fenv f = (\vec y, S_f)
\\ Q_f \ge 0
\\ Q'_f \ge 0
\\\\ B; (\Gamma_f', Q'_f); (\Gamma_f[\Vargs/\vec y], Q_f[\Vargs/ \vec y] ) \vdash S_f \dashv (\Gamma', Q')
}
{ (\Gamma_f, Q_f, \Gamma_f', Q_f') \in \Delta(f) }
%
\and \Rule{Q:Weak}
{  \!\! \Gamma_1 \models \Gamma_2
\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ \Gamma'_2 \models \Gamma'_1
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy} {\in} \N \, . \, \Gamma \models l_{xy} \le |\inter x y| \}
\\ \mathcal U = \{ xy \mid \exists u_{xy}{\in} \N \, . \, \Gamma \models |\inter x y| \le u_{xy} \}
\\ \forall i.\, p_i, r_i \in \Qplusz
\\\\ \forall i \in \mathcal U.\, q'_i \ge q_i - r_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U {\cup} \mathcal L {\cup} \{0\} .\, q'_i \ge q_i
\\ q'_0 \geq q_0 {\textstyle + \sum_{i \in \mathcal U} u_i r_i - \sum_{i \in \mathcal L} l_i p_i}
}
{ Q' \succeq_\Gamma Q }
\end{mathpar}
\caption{Inference rules of the quantitative analysis.}
\label{fig:auto}
\end{figure*}
}{}

\ifshort{
                    \jan{XXX THIS FIGURE NEEDS TO BE UPDATED XXX}
}{}

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x - 10 \geq 0)~\{x=x{-}10; tick(5);\}~else~break;}}}
\begin{figure*}[t]
\vspace{-.5cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=1cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2.5cm,rightskip=2.5cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:Dec)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (\cdot, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (x {\ge} 0, P^\text{de})
          }
        }
        { (\cdot, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (\cdot, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (\cdot, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (\cdot, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (\cdot, B^\text{br}); (\cdot,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (\cdot, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (\cdot, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (\cdot, P)
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(x{\ge}0)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced $\toolname$. The constraints are resolved by an off-the-shelf LP solver.}
\label{fig:derivation}
\end{figure*}

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context
used to gather information on interval sizes.  For example,
$P^\text{de} \succeq_{(x{\ge}0)} P^\text{we} \equiv$
% SSS SPACING
%\vspace{-.3cm}
{\small
\begin{align*}
  p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0,10}
  \land p^\text{we}_{x,0} \leq p^\text{de}_{x,0} + u_{x,0} - v_{0,x} \land \;\;
  \\
  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10}
  \bigwedge_{(\alpha, \beta) \neq (0,10), (x,0)} p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;.
\end{align*}
}
% SSS SPACING
%\vspace{-.8cm}

@@

% SSS SPACING
%\vspace{-.24cm}

@@

\iffull{}{The main parts of the soundness proof are formalized with
  Coq and available for download. The full definitions of the cost
  semantics and the quantitative Hoare logic, and more details on the
  soundness proof can be found in the extended version of this
  article.}

@@

The small-step semantics is standard, except that it tracks
the resource consumption of a program.  The semantics is parametric in
the resource of interest for the user of our system.  We achieve this
independence by using a tick instruction that allows the programmer
to specify how much and when resource is consumed.  If the resource
usage is bound to the programming language (e.g. stack usage, number
loop iterations and function calls), the tick statements can
be inserted automatically by a pre-processing phase.  Resources can be
released by using a negative tick.
Two sample rules of the semantics for update and tick follow.

\begin{mathpar}
\Rule{U}
{ \state' = \state[x \mapsto \sem{e}_\state] }
{ (\state, x \gets e, \cont, \cost)
\rightarrow\\\\
(\state', \code{skip}, \cont, \cost)
}
\and
\Rule[leftskip=0.3cm,rightskip=0.2cm]{T}
{ }
{ (\state, \code{tick}(n), \cont, \cost)
\rightarrow\\\\
(\state, \code{skip}, \cont, \cost {-} n)
}
\end{mathpar}
The rules have as implicit side condition that $\cost$ is
non-negative.  This makes it possible to detect a resource crash as a
stuck configuration where $\cost<0$.

@@

We define quantitative Hoare triples
as $B;R\vdash\htriple Q S {Q'}$ where $B$, $R$, $Q$, and $Q'$ are
maps from program states to an element of $\Q\cup\{\infty\}$ that represents
an amount of resources available.  The assertions $B$ and $R$ are
postconditions for the case in which the block $S$ exits by a \code{break}
or \code{return} statement.  Additionally, $R$ depends on the return
value of the current function.  The meaning of the triple $\htriple Q
S {Q'}$ is as follows: If $S$ is executed with starting state $\state$,
the empty continuation, $Q'(\state) \ge 0$ for all state $\sigma$, and at least $Q(\state)$ resource units available
then the evaluation does not run out of resources and there are at
least $Q'(\state')$ resources left if the evaluation terminates in state
$\state'$.  The logic rules are similar to the ones in previous
work.

@@

The translation of an assertion $(\Gamma, Q)$ in the automatic
analysis is defined by
$$
  \tr(\Gamma, Q) := \lambda\state .\, (\Gamma(\state)) + \Phi_Q(\state),
$$
where we write $\Phi_Q$ for the unique linear potential function
defined by the quantitative annotation $Q$.  The logical context $\Gamma$
is implicitly lifted to a quantitative assertion by mapping a state
$\state$ to 0 if $\Gamma(\state)$ holds and to $\infty$ otherwise.
\iffull{We also need to
translate the assumptions in function contexts of the automatic
analysis.  We define $\tr(\Gamma_f, Q_f, \Gamma'_f, Q'_f) :=$
%
\begin{multline*}
\forall z \, \vec v \, v . \,
(
  \lambda \vec v \, \state .\, \tr(\Gamma_f, Q_f)(\state[\Vargs {\mapsto} \vec v]), \\
  \lambda v \, \state .\, \tr(\Gamma'_f, Q'_f)(\state[\Vret {\mapsto} v])
  )
\end{multline*}}{}%
%
\iffull{These definitions let}{This lets} us translate the judgement
$J := B; R; P \vdash S \dashv P'$\iffull{ in the context $\Delta$ by $\tr(J) :=$}{ by}
\iffull{$$
  (\lambda f .\, \tr(\Delta(f))); \tr(B); \tr(R) \vdash
    \htriple{\tr(P)}{S}{\tr(P')}.
$$}{%
$$
  \tr(J) := \tr(B); \tr(R) \vdash \htriple{\tr(P)} S {\tr(P')}.
$$}%
The soundness of the automatic analysis can now be stated
formally with the following theorem.
%
\begin{theorem}[Soundness of the automatic analysis]
  If $J$ is a judgement derived by the automatic analysis,
  then $\tr(J)$ is a quantitative Hoare triple derivable
  in the quantitative logic.
\end{theorem}
%
\noindent
The proof of this theorem is constructive and maps each rule
of the automatic analysis directly to its counterpart in the
quantitative logic.  The most tricky parts are the translations of the
rules for increments and decrements and the rule {\sc Q:Weak} for
weakening because they make essential use of the algebraic properties
of our linear potential functions.
\iffull{ For instance we prove the following lemma to show the soundness
of the translation of {\sc Q:Weak}.
